Wigglyhedra

### 2025-01-06 MZ/answerReferee.tex

\item \textsl{\color{gray} Complexes associated to curves and extensions in the linear heart: This is perhaps more pedantic, but I think it is nonetheless important. I believe that under your definition, every curve, and in particular every wiggly arc, is only associated with a complex up to overall shifts in internal (path length) and homological gradings. While I understood that many of the claims involving these objects should be implicitly considered “up to both gradings shifts”, I think the claims would benefit from a more precise wording. For example, Proposition 94 as it currently is can be confusing and is falsifiable. For concreteness take $n = 1$ and the arcs $\alpha := (0, 1, \emptyset, \emptyset)$ and $\alpha' := (1, 2, \emptyset, \emptyset)$, which are not compatible (non-pointed). If I were to pick $X := P_1$ and $X' := P_2$, then $\mathrm{Ext}^1(X, X') = 0$ in $A_2$, which contradicts Proposition 94. Of course there are ``better'' choices of $X$ and $X'$, which would make the statement true. One way to correctly state it would be, e.g. ``$\mathrm{Ext}^1(X,X') = 0$ for all representatives $X$ and $X'$ of $\alpha$ and $\alpha'$ respectively''. Similar issue appears in Proposition 97, where as is currently stated, there may or may not be an extension depending on the choice of $X, X', Y, Y'$.} \\
  Indeed, the correct functor to use here is not \(\mathrm{Ext}^1\), but rather a direct sum over \(\mathrm{Ext}^1\)s from \(X\) to all level-zero shifts of \(X'\).
  This has been corrected along the lines you suggested, and the new functor is denoted \(\mathbb{E}^1\).

@@

\item \textsl{\color{gray} pg.\,17 (C2): I assume this is a typo: $g(\alpha)$ should be $\hat{g}(\alpha)$.} \\
Thanks! Fixed.

\item \textsl{\color{gray} pg.\,17 (C3): Unintended capitalisation: should be $g(T)$.} \\
Thanks! Fixed.

\item \textsl{\color{gray} pg.\,18 (C1): I think $1$ and $-1$ should be $2$ and $-2$ respectively.} \\
Thanks! Fixed.

\item \textsl{\color{gray} pg.\,26 (C1): The definition of zigzag algebra as it is would not be correct for $m = 2$; adding in the relation (superfluous for $m > 2$) ``all paths of length $3$ and above are killed'' should suffice.} \\
Thanks! Fixed.

\item \textsl{\color{gray} pg.\,27 (C1): I'd add the adjective ``full''.} \\
Thanks! Fixed and updated to full and isomorphism-closed.

@@

\item \textsl{\color{gray} pg.\,29 (C1): Given your setting, I assume you meant ``derived category'' instead?} \\
TODO

\item \textsl{\color{gray} pg.\,30 (C1): Is ``irrelevant'' meant to be ``external''? I’m not sure where ``irrelevant wiggly arc'' was defined.} \\
Thanks! Fixed.

\item \textsl{\color{gray} pg.\,31 (C1)*: As mentioned in the previous section.} \\
Thanks! Fixed as above.

\item \textsl{\color{gray} pg.\,31 (C2): ``up to equivalence'' should probably also include ``up to scalar multiplication''; same for ``unique element''.} \\
Thanks! Fixed.

\item \textsl{\color{gray} pg.\,32 (C1)*: As mentioned in the previous section, some sort of ``up to appropriate shift in gradings'' should be included.} \\
Thanks! Fixed.

@@

\end{itemize}

Finally, we have also added Figure 15 providing a stereographic projection of the wiggly fan, and we have corrected a few additional typos.

### 2025-01-06 draft/wigglyhedra.tex
Added most of the edits.

\begin{definition}
The \defn{zigzag algebra} \(A_m\) corresponding to the quiver \(Q_m\) depicted in~\cref{fig:am-quiver} is the quotient of \(kQ_m\) by the relations \((i-1, i, i+1) = 0 = (i+1, i, i-1)\) and \((i, i-1, i) = (i, i+1, i)\) for all~\(1 < i < m\).
\end{definition}

@@

\begin{definition}
The category \(\mathcal{C}_m\) is defined to be the smallest triangulated subcategory of \(D^b(A_m\text{-mod})\) containing all the objects \(P_i\langle j \rangle\).
\end{definition}

@@

\begin{proposition}
\label{prop:hn-filtration}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider its cohomology filtration
  \[
    \begin{tikzcd}[column sep=0.3cm]
      0 \arrow{rr} && X_{-\ell} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^\ell(X)[-\ell] \arrow[dashed]{ul}{+1} && H^{\ell+1}(X)[-\ell-1] \arrow[dashed]{ul}{+1} & \cdots & H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]
  Consider the set of all possible indecomposable summands of the cohomology objects \(\{H^j(X)\}\) taken without multiplicity.
  Let \(T\) be corresponding set of wiggly arcs, as explained in~\cref{prop:cohomology-pieces}.~Then
  \begin{enumerate}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation,
  \item \(T\) contains at most one of the two irrelevant wiggly arcs.
  \end{enumerate}
\end{proposition}

@@

Recall that two wiggly arcs \(\alpha\) and \(\alpha'\) are said to be compatible (as in~\cref{def:compatible}) if they are non-crossing and pointed.
The incompatibility degree of \(\alpha\) and \(\alpha'\) (as in~\cref{def:incompatibilityDegree}) is \(1\) if \(\alpha\) and \(\alpha'\) are non pointed, and the number of crossings of \(\alpha\) and \(\alpha'\) otherwise.
\begin{proposition}
\label{prop:ext1}
Let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to two wiggly arcs \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible if and only if \(\Ext^1(X, X') = 0\).
  More generally, the incompatibility degree of \(\alpha\) and \(\alpha'\) is the dimension of \(\Ext^1(X,X')\).
\end{proposition}

@@

With our path-length grading, note that a morphism in \(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1])\) has level \(s_1 - s_2\); that is, it takes an object \(X\) of level \(0\) to the object \(X'\langle -s_2 \rangle[s_1]\) of level \(s_1 - s_2\).
  We can thus obtain the dimension of all morphisms of level \(\ell\) by setting \(q_1 = q\) and \(q_2 = q^{-1}\) and taking the coefficient of \(q^{\ell}\) in the expression~\eqref{eq:bigraded-homs}.
    The dimension of \(\Ext^1(X,X')\), is therefore the coefficient of \(q\) under this specialisation.

It can be checked that under the path-length grading, the space of morphisms arising from an intersection of curves of ``crossing'' type as well as ``non pointed'' type contains exactly one morphism (up to equivalence) of level one; that is, a unique element of \(\Ext^1\).
Morphisms arising from other intersections do not contribute to \(\Ext^1\).
This calculation proves the proposition.
\end{proof}

@@

\begin{remark}
  As stated,~\cref{prop:ext1} in combination with~\cref{prop:hn-filtration} is not expected to be true in greater generality.
  For instance, for the zigzag algebra of Dynkin type~\(D_4\), one can find two spherical objects with a non-trivial \(\Ext^1\) that nevertheless appear together in the cohomology filtration of another spherical object.
  However, we expect to see similar constraints on compatibility, which we hope to turn to in forthcoming work.
\end{remark}

@@

Let \(X\), \(X'\), \(Y\) and~\(Y'\) be spherical objects in the standard heart that correspond to \(\alpha\), \(\alpha'\), \(\beta\) and~\(\beta'\) respectively.
First, note by~\cref{prop:ext1} that \(X\) and \(X'\) are exchangeable if and only if there is a unique non-trivial extension between the corresponding objects of the standard heart.
Next, let us understand~\(Y\) and \(Y'\) categorically.

\begin{proposition}\label{prop:extensions}
For \(\alpha\), \(\alpha'\), \(\beta\), \(\beta'\), and \(X\), \(X'\), \(Y\), \(Y'\) as just discussed,
\begin{itemize}
\item if \(\alpha\) and \(\alpha'\) are not pointed, then we have extensions
  \[0 \to X' \to Y \to X \to 0 \quad \text{and}\quad 0 \to X \to Y' \to X' \to 0,\]
\item if \(\alpha\) and \(\alpha'\) are crossing, then we have an extension
  \[0 \to X' \to (Y \oplus Y') \to X \to 0.\]
\end{itemize}
\end{proposition}

@@

\begin{question}
  Suppose we have a collection of spherical objects in the standard heart \(\mathcal{A}_m\) that satisfy the compatibility condition of~\cref{prop:ext1}.
  That is, the \(\Ext^1\) between any pair of objects in this collection is zero.
  \begin{enumerate}
  \item Can we prove properties about a maximal such collection without reference to curves?
  \item In particular, can we understand the combinatorics of maximal such collections without reference to curves?
  \item Given a maximal compatible collection of objects, can we always find a larger spherical object such that the factors of the cohomology filtration of that object are precisely shifts of the objects in the given collection?
  \end{enumerate}
\end{question}

>>>>

\begin{definition}
The \defn{zigzag algebra} \(A_m\) corresponding to the quiver \(Q_m\) depicted in~\cref{fig:am-quiver} is the quotient of \(kQ_m\) by all paths of length at least \(3\), together with the relations \((i-1, i, i+1) = 0 = (i+1, i, i-1)\) and \((i, i-1, i) = (i, i+1, i)\) for all~\(1 < i < m\).
\end{definition}

@@

\begin{definition}
The category \(\mathcal{C}_m\) is defined to be the smallest full, isomorphism-closed triangulated subcategory of \(D^b(A_m\text{-mod})\) containing all the objects \(P_i\langle j \rangle\).
\end{definition}

@@

\begin{proposition}
\label{prop:hn-filtration}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider its cohomology filtration
  \[
    \begin{tikzcd}[column sep=0.3cm]
      0 \arrow{rr} && X_{-\ell} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^\ell(X)[-\ell] \arrow[dashed]{ul}{+1} && H^{\ell+1}(X)[-\ell-1] \arrow[dashed]{ul}{+1} & \cdots & H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]
  Consider the set of all possible indecomposable summands of the cohomology objects \(\{H^j(X)\}\) taken without multiplicity.
  Let \(T\) be corresponding set of wiggly arcs, as explained in~\cref{prop:cohomology-pieces}.~Then
  \begin{enumerate}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation,
  \item \(T\) contains at most one of the two external wiggly arcs.
  \end{enumerate}
\end{proposition}

@@

For two objects \(X\) and \(X'\) of \(\mathcal{C}_m\), let \(\mathbb{E}^\ell(X,X')\) denote the direct sum over all level-\(\ell\) morphisms from \(X\) to \(X'\).
That is,
\[\mathbb{E}^\ell(X,X') = \bigoplus_{g + h = \ell}\Hom^{g,h}(X,X').\]

Recall that two wiggly arcs \(\alpha\) and \(\alpha'\) are said to be compatible (as in~\cref{def:compatible}) if they are non-crossing and pointed.
The incompatibility degree of \(\alpha\) and \(\alpha'\) (as in~\cref{def:incompatibilityDegree}) is \(1\) if \(\alpha\) and \(\alpha'\) are non pointed, and the number of crossings of \(\alpha\) and \(\alpha'\) otherwise.
\begin{proposition}
\label{prop:ext1}
Let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to two wiggly arcs \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible if and only if \(\mathbb{E}^1(X, X') = 0\).
  More generally, the incompatibility degree of \(\alpha\) and \(\alpha'\) is the dimension of \(\mathbb{E}^1(X,X')\).
\end{proposition}

@@

With our path-length grading, note that a morphism in \(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1])\) has level \(s_1 - s_2\); that is, it takes an object \(X\) of level \(0\) to the object \(X'\langle -s_2 \rangle[s_1]\) of level \(s_1 - s_2\).
  We can thus obtain the dimension of all morphisms of level \(\ell\) by setting \(q_1 = q\) and \(q_2 = q^{-1}\) and taking the coefficient of \(q^{\ell}\) in the expression~\eqref{eq:bigraded-homs}.
    The dimension of \(\mathbb{E}^1(X,X')\), is therefore the coefficient of \(q\) under this specialisation.

It can be checked that under the path-length grading, the space of morphisms arising from an intersection of curves of ``crossing'' type as well as ``non pointed'' type contains exactly one morphism (up to equivalence and scalar multiplication) of level one; that is, a unique element (up to scalar multiplication) of \(\mathbb{E}^1\).
Morphisms arising from other intersections do not contribute to \(\mathbb{E}^1\).
This calculation proves the proposition.
\end{proof}

@@

\begin{remark}
  As stated,~\cref{prop:ext1} in combination with~\cref{prop:hn-filtration} is not expected to be true in greater generality.
  For instance, for the zigzag algebra of Dynkin type~\(D_4\), one can find two spherical objects with a non-trivial \(\mathbb{E}^1\) that nevertheless appear together in the cohomology filtration of another spherical object.
  However, we expect to see similar constraints on compatibility, which we hope to turn to in forthcoming work.
\end{remark}

@@

Let \(X\), \(X'\), \(Y\) and~\(Y'\) be spherical objects in the standard heart that correspond to \(\alpha\), \(\alpha'\), \(\beta\) and~\(\beta'\) respectively.
First, note by~\cref{prop:ext1} that \(X\) and \(X'\) are exchangeable if and only if there is a unique non-trivial \(\mathbb{E}^1\) between the corresponding objects of the standard heart.
Next, let us understand~\(Y\) and \(Y'\) categorically.

\begin{proposition}\label{prop:extensions}
For \(\alpha\), \(\alpha'\), \(\beta\), \(\beta'\), and \(X\), \(X'\), \(Y\), \(Y'\) as just discussed,
\begin{itemize}
\item if \(\alpha\) and \(\alpha'\) are not pointed, then for some \(g\) and \(g'\), we have extensions
  \[0 \to X' \to Y \to X\langle g \rangle[-g] \to 0 \quad \text{and}\quad 0 \to X \to Y' \to X'\langle g' \rangle[-g'] \to 0;\]
\item if \(\alpha\) and \(\alpha'\) are crossing, then for some \(g\), we have an extension
  \[0 \to X' \to (Y \oplus Y') \to X\langle g \rangle[-g]\to 0.\]
\end{itemize}
\end{proposition}

@@

\begin{question}
  Suppose we have a collection of spherical objects in the standard heart \(\mathcal{A}_m\) that satisfy the compatibility condition of~\cref{prop:ext1}.
  That is, the \(\mathbb{E}^1\) between any pair of objects in this collection is zero.
  \begin{enumerate}
  \item Can we prove properties about a maximal such collection without reference to curves?
  \item In particular, can we understand the combinatorics of maximal such collections without reference to curves?
  \item Given a maximal compatible collection of objects, can we always find a larger spherical object such that the factors of the cohomology filtration of that object are precisely shifts of the objects in the given collection?
  \end{enumerate}
\end{question}
